---
title: "The Legacy of the European Post-Master in Urbanism at TU Delft: A Text Mining Approach"
author:
  - Jane Doe:
      correspondence: "yes"
      email: janedoe@fosg.org
      orcid: 0000-0003-1689-0557
      institute: tud
  - John Q. Doe:
      email: johndoe@fosg.org
      orcid: 0000-0003-1689-0558
      institute: tud
institute:
  - tud:
      name: Delft University of Technology
      address: Julianalaan 134, 2628 BL, Delft, Zuid-Holland, The Netherlands
title-block-published: "Last updated"  
date: now
date-format: long
format: 
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: true
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  urbanism; education; post-master; text mining
highlights: |
  These are the highlights. 
---

<!-- With the following code you can access and display values from the yml header above. -->

Keywords: `r rmarkdown::metadata$keywords`

Highlights: `r rmarkdown::metadata$highlights`

<!-- The actual document text starts here: -->

```{r}
#| label: setup
#| echo: false
#| warning: false

# Load packages ----
## List packages to be loaded
packages <- c(
  "here",                                        # Managing paths
  "tidyverse",                                   # Data manipulation and visualisation
  "tidytext", "SnowballC", "tm", "topicmodels",  # Text processing and analysis
  "tidygeocoder",                                # Working with spatial data
  "igraph", "ggraph"                             # Network analysis and visualsation
)

## Load required packages, install them if needed
for (package in packages) {
  if (!require(package, character.only = TRUE)) {
    install.packages(package,  "http://cran.us.r-project.org")
  }
  library(package, character.only = TRUE)
}

# library(here)
# library(tidyverse)
# library(tidytext)
# library(SnowballC)
# library(tm)
# library(topicmodels)
# library(tidygeocoder)
# library(igraph)
# library(ggraph)



# Load analysis functions ----
devtools::load_all(".")


# Read thesis metadata ----
data_path <- here("analysis", "data", "raw_data")
pdf_names <- dir(data_path, pattern = "*.pdf")

## Read metadata for all theses
all_theses <-
  readxl::read_xlsx(path = here(data_path, "theses-all.xlsx"))

## Read only metadata for theses with PDF available
emu_theses <- 
  read_csv(here(data_path, "theses-metadata.csv")) %>%
  mutate(text = "") %>%
  filter(!is.na(pdf_via))

pdf_paths <- here(data_path, emu_theses$file_name)
```

```{r}
#| label: extract-pdf-text
#| echo: false

# Extract text from PDFs
# On the first run, this function might take extra time as a Python environment needs to be set up

emu_theses$text <- convert_pdf_text(pdf_paths)
```

```{r}
#| label: clean-text
#| echo: false

emu_theses$text_clean <- emu_theses$text |>
  clean_basic() 

```

# Introduction

The research presented in this paper was prompted by the closure of the European post-Master of Urbanism (EMU) of the Department of Urbanism at the Faculty of Architecture and the Built Environment, TU Delft. The EMU was an advanced master ran jointly by TU Delft, KU Leuven, UPC Barcelona and Università IUAV di Venezia. <!-- Complete paragraph about the program -->

Distinctive features of the program:  
- as a postmaster, it only accepted applicants with a prior MSc degree and experience in practice
- provides a bridge towards PhD-level research
- all students had an exchange semester at one of the other three program universities

In order to describe the legacy of the EMU program, including the distinctive features of its didactic approach, this paper aims to reveal the main topics taught in it and how those topics had evolved through the years of the program. To that end, we employed a text mining approach in which we analysed its output: `r nrow(all_theses)` theses with an average of `r mean(str_length(emu_theses$text_clean), na.rm = TRUE)` words produced over the years for the duration of the program between `r paste0(min(all_theses$grad_year, na.rm = TRUE), "-", max(all_theses$grad_year, na.rm = TRUE))`. The first year of the program was not represented, as PDF files were only available from 2008 onward.

# Methods

Research questions:

-   What were the main topics addressed in the EMU program?

-   How did those topics evolve throughout the duration of the program between 2007-2021?

-   To what extent were the topics and evolution thereof influenced by the assignments given to the students throughout their studies leading up to their thesis?

-   To what extent were the topics and evolution thereof influenced by the exchange semester followed by the students?

## Data collection

The theses are available in PDF format with complex layouts typical to an urbanism project in which text and various types of graphics are combined. With the exception of a bilingual English-Spanish thesis, all thesis were written in English.

A total number of `r nrow(emu_theses)` out of `r nrow(all_theses)` were analysed.

```{r}
#| label: fig-map-theses
#| warning: false 
#| echo: false
#| fig-cap: "Location of EMU theses included in the analysis"

# Add latitude and longitude information to theses data frame
emu_theses <- geocode_thesis_locations(emu_theses, location = NULL)

# Show thesis locations on world map
visualize_thesis_locations(emu_theses)
```

Figure @fig-map-theses shows the geographic spread of the theses included in the analysis.

## Data analysis

[@silge2017a] [@hvitfeldt2021a]

### Tokenisation

Word-level tokenisation was applied with the `unnest_tokens()` function of the `tidytext` package. This was preferred over other tokenisers as it produced results that work seamlessly with other `tidyverse` tools used in the analysis.

### Stop words

For analyses in which word-level tokenisation was chosen, stop words, i.e., common words or words that do not add much meaning, were removed in the data preparation stage. Global, subject-specific and document stop words [@hvitfeldt2021a] were differentiated. As all theses were written in English, the pre-made "snowball" lexicon was used to remove global stop words. <!-- We tried to create a custom global stop word list on the entire corpus, by removing high-frequency words, starting with 20 words and increasing by 10 until reaching words that are not appropriate as stop words, but no significant imoprvement was observed in the result. --> Subject-specific stop words, we manually constructed a list based on our domain knowledge. Words that are commonly used to structure a theses, such as "preface", "contents", and "introduction", were added as subject-specific stop words as well. Finally, titles and author names were removed as document stop words on document-level.

```{r}
#| label: stop-words

emu_theses$text_clean <- emu_theses$text_clean |>
  str_remove_all(find_meta_stopwords(emu_theses)) |> 
  str_remove_all(urbanism_stopwords(
    add_stopwords = c("emu", "european postgraduate master in urbanism"))) |>
  str_remove_all(thesis_stopwords(add_stopwords = c("advisor", "prof")))

emu_theses_words <- emu_theses |> 
  select(title, text_clean) |> 
  unnest_tokens(output = word, input = text_clean) |>   # remove punctuation, convert to lowercase, separate all words
  anti_join(tidytext::stop_words, by = "word")
```

Stemming is a pre-processing step that needs to be carefully thought through, as it might either degrade topic modeling or produce no meaningful results [@hvitfeldt2021a].

```{r}
#| label: stemming

emu_theses_words %>% 
  mutate(stem = wordStem(words = word, 
                         language = "porter")) %>%  # Implement the Porter stemming algorithm for English provided by the SnowballC package
  count(stem, sort = TRUE)
```

## Topic model

Similar to topic clustering on numeric data, topic modeling is a method of unsupervised classification of topics found in a group of documents [@silge2017].

# Results

```{r}
# Which are the top n most frequently used words in each thesis?
get_top_words_per_document(data = emu_theses_words, 
                           top_n = 3, 
                           title_col = "title", 
                           word_col = "word")
```

```{r}
# Which are the top n most frequently used words in all theses`?
get_top_words_per_corpus(data = emu_theses_words, 
                         top_n = 5, 
                         word_col = "word")
```

```{r}
# Which are the top n most frequently used words in all theses`?
emu_theses_top_20 <- get_top_words_per_corpus(data = emu_theses_words, 
                         top_n = 20, 
                         word_col = "word")
```

<!-- ```{r} -->

<!-- # Which theses contain "fabric" in the top 20 most frequently used words? -->

<!-- emu_theses_top_20 %>%  -->

<!--   filter(word == "fabric") -->

<!-- emu_theses_top_20 %>% -->

<!--   slice(1:40) %>%  -->

<!--   ggplot(aes(x = reorder(word, n), y = n, fill = title)) + -->

<!--   geom_col() + -->

<!--   xlab(NULL) + -->

<!--   coord_flip() + -->

<!--   labs(y = "Count", -->

<!--        x = "Unique words", -->

<!--        title = "Count of unique words found in thesis", -->

<!--        subtitle = "Stop words were removed from the list") + -->

<!--   facet_wrap( ~ title, scales = "free", ncol = 2) -->

<!-- ``` -->

```{r}
#| label: ngrams

## TO DO: generalise bigrams and trigrams into ngrams function

## Bigrams
emu_theses_bigrams <- emu_theses |> 
  unnest_tokens(bigram, text, token = "ngrams", n = 2) |> 
  separate(bigram, into = c("first", "second"), sep = " ", remove = FALSE) |> 
  anti_join(stop_words, by = c("first" = "word")) |> 
  anti_join(stop_words, by = c("second" = "word")) |> 
  filter(str_detect(first, "[a-z]") & 
         str_detect(second, "[a-z]"))

bigram_counts <- emu_theses_bigrams |> 
  count(first, second, sort = TRUE)

bigram_counts

## Trigrams
emu_theses_trigrams <- emu_theses |> 
  unnest_tokens(trigram, text, token = "ngrams", n = 3) |> 
  separate(trigram, into = c("first", "second", "third"), sep = " ", remove = FALSE) |> 
  filter(!first %in% stop_words$word,
         !second %in% stop_words$word,
         !third %in% stop_words$word) |>  
  filter(str_detect(first, "[a-z]"), 
         str_detect(second, "[a-z]"),
         str_detect(third, "[a-z]"))

trigram_counts <- emu_theses_trigrams |> 
  count(first, second, sort = TRUE)

trigram_counts

## Bigram tf-idf
bigram_tf_idf <- emu_theses_bigrams |> 
  count(title, bigram) |> 
  bind_tf_idf(bigram, title, n) |> 
  arrange(desc(tf_idf))

bigram_tf_idf

## Trigram tf-idf
trigram_tf_idf <- emu_theses_trigrams |> 
  count(title, trigram) |> 
  bind_tf_idf(trigram, title, n) |> 
  arrange(desc(tf_idf))

trigram_tf_idf

## Show network of bigrams
bigram_graph <- bigram_counts |> 
  filter(n > 100) |> 
  graph_from_data_frame()

bigram_graph

set.seed(2023)

a <- grid::arrow(type = "closed", length = unit(.15, "inches"))
  
ggraph(bigram_graph, layout = "fr") +
  geom_edge_link(aes(edge_alpha = n), show.legend = FALSE, arrow = a) +
  geom_node_point(color = "lightblue", size = 5) +
  geom_node_text(aes(label = name), vjust = 1, hjust = 1) +
  theme_void()

## Higgh frequency bigrams ----

emu_theses_bigrams_top <- emu_theses_bigrams %>%
  group_by(title) %>%
  count(bigram, sort = TRUE) %>%
  slice_max(n, n = 20) %>%        ### top_n is superseded; better use slice_max instead
  # filter(n > 50)                ### filter words with a minimum count
  ungroup() %>% 
  mutate(bigram = reorder(bigram, n))

emu_theses_bigrams_top_all <- emu_theses_bigrams %>%
  count(bigram, sort = TRUE) %>%
  slice_max(n, n = 20) %>%        ### top_n is superseded; better use slice_max instead
  # filter(n > 50)                ### filter words with a minimum count
  mutate(bigram = reorder(bigram, n))

emu_theses_bigrams_top %>%
  filter(title == unique(title)[1:6]) |> 
  ggplot(aes(x = bigram, y = n, fill = title)) +
  geom_col() +
  xlab(NULL) +
  coord_flip() +
  labs(y = "Count",
       x = "Unique words",
       title = "Count of unique words found in thesis",
       subtitle = "Stop words were removed from the list") +
  facet_wrap( ~ title, scales = "free", ncol = 2) +
  theme(legend.position = "none") 
```

```{r}
#| label: tf-idf

emu_theses_words

emu_theses_words_count <- emu_theses_words |> 
  count(title, word, sort = TRUE) |> 
  ungroup()

emu_theses_words_total <- emu_theses_words_count |> 
  group_by(title) |> 
  summarise(total = sum(n))

emu_theses_words_count <- left_join(emu_theses_words_count, emu_theses_words_total)

# emu_theses_words_count |> 
#   group_by(title) |> 
#   summarise(total = unique(tot

# Preview term frequency distributions in a subset of theses
ggplot(emu_theses_words_count |> 
         filter(title == unique(title)[1:6]),
       aes(n/total, fill = title)) +
  geom_histogram(show.legend = FALSE, binwidth = 0.0001) +
  xlim(NA, 0.0009) +
  facet_wrap(~ title, ncol = 3, scales = "free_y")

# Examine Zipf's law for the EMU theses
## Calculate rank and term frequency
freq_by_rank <- emu_theses_words_count |> 
  group_by(title) |> 
  mutate(rank = row_number(),
         `term frequency` = n/total)

## Visualise Zipf's law
freq_by_rank |> 
  ggplot(aes(rank, `term frequency`, color = title)) +
  geom_line(size = 1.1, alpha = 0.8, show.legend = FALSE) +
  scale_x_log10() +
  scale_y_log10()

rank_subset <- freq_by_rank |> 
  filter(rank < 500,
         rank > 10)

lm(log10(`term frequency`) ~ log10(rank), data =  rank_subset)

freq_by_rank |> 
  ggplot(aes(rank, `term frequency`, color = title)) +
  geom_line(size = 1.1, alpha = 0.8, show.legend = FALSE) +
  geom_abline(intercept = -1.44, slope = -0.7, color = "gray50", linetype = 2) +
  scale_x_log10() +
  scale_y_log10()

## Calculate tf-idf
emu_theses_words_count <- emu_theses_words_count |> 
  bind_tf_idf(word, title, n)
emu_theses_words_count

emu_theses_words_count |> 
  select(-total) |> 
  arrange(desc(tf_idf))

## Visualise the top n words with the highest tf-idf values for each of a subset of theses
emu_theses_words_count |> 
  filter(title == unique(title)[1:6]) |> 
  arrange(desc(tf_idf)) |> 
  mutate(word = factor(word, levels = rev(unique(word)))) |> 
  group_by(title) |> 
  top_n(15) |> 
  ungroup() |> 
  ggplot(aes(word, tf_idf, fill = title)) +
  geom_col(show.legend = FALSE) +
  labs(x = NULL, y = "tf-idf") +
  facet_wrap(~ title,  ncol = 3, scales = "free") +
  coord_flip()
```

The `tf-idf` statistic shows high values for the names of thesis locations that tend to be high-frequency thesis-specific words. As such, it shows words that are the most important to one document in a collection of documents [@silge2017].

# Discussion

# Conclusion

# Acknowledgements

<!-- The following line inserts a page break  -->

\newpage

# References

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

::: {#refs}
:::

\newpage

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r}
#| label: colophon
#| cache: false

# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```
